GHRVis:

### 2022-02-16 paper/section/main.tex
draft: update main section

\subsection{Node Overlap Removal Algorithm}

% \begin{noindent}
\begin{algorithm}[hbt!]
    \caption{An overview of all steps in \software.}\label{alg:overview}
    \textbf{Variables:} \\
    $NodeList \gets$ a list of nodes representing CCGs \\
    $Node_{size}\gets$ the current size of all nodes \\
    $Node_{inc} \gets$ the increment node size of each iteration \\
    $Node_{max} \gets$ the maximum size of a node \\

    \begin{algorithmic}[1]
        \While{$ Node_{size} < Node_{max} $}
            \State RemoveOverlap ($ NodeList $)
            \State $Node_{size} \gets Node_{size} + Node_{inc}$
        \EndWhile
    \end{algorithmic}
\end{algorithm}

\begin{algorithm}[hbt!]
    \caption{A function to remove node overlaps.}\label{alg:remove overlap}
    \textbf{Variables:} \\
    $Node_{stale} \gets$ the maximum number of iterations before a stalemate \\
    \begin{algorithmic}[1]
        \Function{RemoveOverlap}{$ NodeList $}
        \State $ recurse \gets False$
        \State $ nodeList_{ora} \gets $ RunOverlapRemoval ($ NodeList $)
            \For{$ node^i_{ora} \in nodeList_{ora} $}
                \If {MoveNode ($ Node^i, node^i_{ora} $)}
                    \State $ cross \gets $ CheckRiverCrossing ($ Node^i, node^i_{ora} $)
                    \If{$ cross = True $}
                        \State $recurse \gets True$
                        \State $ Node^i.stale \gets Node^i.stale + 1 $
                        \If{$ Node^i.stale < Node_{stale} $}
                            \State MoveNode ($ Node^i $)
                        \Else
                            \State DeriveCorridor ($ Node^i, node^i_{ora} $)
                        \EndIf
                    \EndIf
                \EndIf
            \EndFor
            \If{$ recurse = True $}
                \State RemoveOverlap ($ NodeList $)
            \EndIf
        \EndFunction
    \end{algorithmic}
\end{algorithm}

\begin{algorithm}[hbt!]
    \caption{A function to update a node's position.}\label{alg:move position}
    \begin{algorithmic}[1]
        \Function{MoveNode}{$ Node, Node_{ora} $}
        \If{$ Node.x = Node_{ora}.x $ and $ Node.y = Node_{ora}.y $}
            \State \Return{$ False $}
        \EndIf

@@

        \If{$ Node_{ora}$ is $ undefined $}
            \State $previous \gets Node.history.pop() $
            \State $ x \gets previous.x,~y \gets previous.y $
        \Else
            \State $ x \gets Node_{ora}.x,~y \gets Node_{ora}.y $
            \State $ Node.history.push(x, y) $
        \EndIf

        \State $ Node.x \gets x $
        \State $ Node.y \gets y $
        \State \Return{$ True $}
        \EndFunction
    \end{algorithmic}
\end{algorithm}

\begin{algorithm}[hbt!]
    \caption{A function to check if a node crosses a river.}\label{alg:check river crossing}
    \textbf{Variables:} \\
    $RiverEdges \gets$ a list of river edges \\

    \begin{algorithmic}[1]
        \Function{CheckRiverCrossing}{$ Node, Node_{ora} $}
            \For{$ edge \in RiverEdges $}
                \State $ node_b \gets $ GenerateBoundary ($ Node, Node_{ora} $)
                \State $ edge_b \gets $ GenerateBoundary ($ edge $)
                \State $ collide \gets $ DetectBoundaryCollision ($ node_b, edge_b $)
                \If{$ collide = True $}
                    \State $ node_l \gets $ GenerateLine ($ Node, Node_{ora} $)
                    \State $ edge_l \gets $ GenerateLine ($ edge $)
                    \State $ intersect \gets $ DetectLineIntersection ($ node_l, edge_l $)
                    \If{$ intersect = True $}
                        \State \Return{$ True $}
                    \EndIf
                \EndIf
            \EndFor
            \State \Return{$ False $}
        \EndFunction
    \end{algorithmic}
\end{algorithm}

\begin{algorithm}[hbt!]
    \caption{A function to derive a corridor to include and move neighboring nodes. On a SVG canvas, the point of origin (0,0) is at the top left corner, with the x-axis extending to the right and the y-axis extending to the bottom, there is no negative axes.}\label{alg:derive corridor}
    \textbf{Variables:} \\
    $C_l \gets$ the length of a corridor \\
    $C_{w} \gets$ the width of a corridor \\
    $Node_{size}\gets$ the current size of all nodes \\
    \begin{algorithmic}[1]
        \Function{DeriveCorridor}{$ Node, Node_{ora} $}
        \State MoveNode ($ Node $)

        \State $ \Delta x \gets Node.x - Node_{ora}.x,~ \Delta y \gets Node.y - Node_{ora}.y $

        \State $ \vec{diff} \gets
            \begin{pmatrix}
                \Delta x \\
                \Delta y
            \end{pmatrix},~ \vec{Node} \gets
            \begin{pmatrix}
                    Node.x \\
                    Node.y
            \end{pmatrix} $

        \State $ \vec{p} \gets
            \vec{Node} + \sqrt{
                C_{l}^2 \cdot (\Delta x + \Delta y))
            } \cdot \vec{diff} $

        \State $ shift \gets
                \sqrt{ 
                    \frac{C_{w}^2}{2}
                },~ \vec{shift_a} \gets \begin{pmatrix}
                    shift \\
                    - shift
                \end{pmatrix},~ \vec{shift_b} \gets \begin{pmatrix}
                    - shift \\
                    +shift
                \end{pmatrix} $

        \State $ corridor \gets
            \begin{bmatrix}
                \vec{Node} + \vec{shift_a} &
                \vec{p}+ \vec{shift_a} \\

                \vec{Node} + \vec{shift_b} &
                \vec{p}+ \vec{shift_b} \\
            \end{bmatrix} $

        \For{$ node^i $ inside $ corridor $}
            \State MoveNode ($ node^i,  $)
        \EndFor

        \EndFunction
    \end{algorithmic}
\end{algorithm}
%\end{noindent}

>>>>

\algoref{alg:UpdateNodePosition} describes the overview of \software. We first load all processed data from \secref{sec:{Data Description}} and apply VPSC \cite{dwyer2006fast} recursively to remove overlaps. We chose VPSC over other node overlap removal algorithms since VPSC is able to provide spread minimization and node movement minimization while maintaining a good level of global shape preservation \cite{chen2020Node}. During each recursion, we check node trajectories (See \algoref{alg:check river crossing}) and move back nodes that crossed a river. A recursion ends when 1) no overlaps are present; 2) no nodes crossed a river. We then increase the node size and repeat the recursion until the entire map canvas is filled.

% \begin{noindent}

\begin{algorithm}[tb!]
    \caption{The procedure to update node positions by removing overlaps and prevent nodes from crossing rivers.}\label{alg:UpdateNodePosition}
    \textbf{Variables:} \\
    $NodeList \gets$ a list of nodes representing CCGs \\
    $Node_{size}\gets$ the current size of all nodes \\
    $Node_{inc} \gets$ the increment node size of each iteration \\
    $Node_{max} \gets$ the maximum size of a node \\
    $Node_{stale} \gets$ the maximum number of iterations before a stalemate \\
    \begin{algorithmic}[1]
        \Procedure{UpdateNodePosition}{}
        \While{$ Node_{size} < Node_{max} $}

        \State $ nodeList_{ora} \gets $ RunVPSC ($ NodeList $)

            \For{$ node^i_{ora} \in nodeList_{ora} $}

                \If {MoveNode ($ Node^i, node^i_{ora} $)}
                    \State $ cross \gets $ \Call{CheckRiverCrossing}{$ Node^i, node^i_{ora} $}

                    \If{$ cross = True $}
                        \State $ Node^i.stale \gets Node^i.stale + 1 $

                        \If{$ Node^i.stale < Node_{stale} $}
                            \State \Call{MoveNode}{$ Node^i $}
                        \Else
                            \State \Call{DeriveCorridor}{$ Node^i, node^i_{ora} $}
                        \EndIf

                    \EndIf

                \EndIf

            \EndFor

        \EndWhile

        \State $Node_{size} \gets Node_{size} + Node_{inc}$

        \EndProcedure
    \end{algorithmic}
\end{algorithm}

\begin{algorithm}[tb!]
    \caption{The procedure to update a node's position.}\label{alg:move position}
    \begin{algorithmic}[1]
        \Procedure{MoveNode}{$ Node, Node_{new} $}
        \If{$ Node.x = Node_{new}.x $ and $ Node.y = Node_{new}.y $}
            \State \Return{$ False $}
        \EndIf

@@

        \If{$ Node_{new}$ is $ undefined $}
            \State $previous \gets Node.history.pop() $
            \State $ x \gets previous.x,~y \gets previous.y $
        \Else
            \State $ x \gets Node_{new}.x,~y \gets Node_{new}.y $
            \State $ Node.history.push(x, y) $
        \EndIf

        \State $ Node.x \gets x $
        \State $ Node.y \gets y $
        \State \Return{$ True $}
        \EndProcedure
    \end{algorithmic}
\end{algorithm}

%\end{noindent}

\subsection{River Cross Checking}

We use rivers as topological boundaries to prevent nodes from crossing them. When a node's position changes, we check if the node's trajectory intersects any segment of a river, see \algoref{alg:check river crossing}. A boundary collision detection is performed to reduce the number of line intersection detections required.

% \begin{noindent}
\begin{algorithm}[tb!]
    \caption{The procedure to check if a node crosses a river.}\label{alg:check river crossing}
    \textbf{Variables:} \\
    $RiverEdges \gets$ a list of river edges formed by two points \\

    \begin{algorithmic}[1]
        \Procedure{CheckRiverCrossing}{$ Node, Node_{ora} $}
            \For{$ edge \in RiverEdges $}
                \State $ node_b \gets $ GenerateBoundary ($ Node, Node_{ora} $)
                \State $ edge_b \gets $ GenerateBoundary ($ edge $)
                \State $ collide \gets $ DetectBoundaryCollision ($ node_b, edge_b $)
                \If{$ collide = True $}
                    \State $ node_l \gets $ GenerateLine ($ Node, Node_{ora} $)
                    \State $ edge_l \gets $ GenerateLine ($ edge $)
                    \State $ intersect \gets $ DetectLineIntersection ($ node_l, edge_l $)
                    \If{$ intersect = True $}
                        \State \Return{$ True $}
                    \EndIf
                \EndIf
            \EndFor
            \State \Return{$ False $}
        \EndProcedure
    \end{algorithmic}
\end{algorithm}
%\end{noindent}

\subsection{Corridor Derivation}

As the VPSC always tries to produce an optimal node layout where node spread and movement are minimized, a node might be repeatedly traveling between two positions due to the lack of available space, creating a stalemate situation for recursions, as shown in \figref{fig:stalemate}. In order to create space for VPSC to generate a new layout, we introduce a user-adjustable parameter, $ Node_{stale} $, to break the stalemate: if a node is traveling between two positions for more than $ Node_{stale} $ iterations, it is considered to be in a stalemate. A stalemate triggers a corridor derivation which moves all nodes within the corridor by $ \mathcal{D} $ set by the user (See \figref{fig:corridor} and \algoref{alg:derive corridor}).

{
\begin{figure}[tb!]
    \centering
    \includegraphics[width=\columnwidth]{figure/stalemate.png}
    \caption{A stalemate situation is when a node repeatedly travels between two positions (A and B) for more than $Node_{stale}$ iterations. }
    \label{fig:stalemate}
\end{figure}
}

{
\begin{figure}[tb!]
    \centering
    \includegraphics[width=\columnwidth]{figure/corridor.png}
    \caption{When a stalemate occurs, a corridor (black rectangular) is derived based on the current (A) and previous (B) positions of the node that crossed a river, as described in \algoref{alg:derive corridor}. All nodes within the corridor are moved by $ \mathcal{D} $ in the direction of $ \vec{BA} $. }
    \label{fig:corridor}
\end{figure}
}

% \begin{noindent}

\begin{algorithm}[tb!]
    \caption{The procedure to derive a corridor to move included nodes. We use a SVG canvas, where the point of origin (0,0) is located at the top left corner, with the x-axis extending to the right and the y-axis extending to the bottom, there is no negative axes.}\label{alg:derive corridor}
    \textbf{Variables:} \\
    $\mathcal{C}_l \gets$ the length of a corridor \\
    $\mathcal{C}_w \gets$ the width of a corridor \\
    $\mathcal{D} \gets$ the distance to move nodes within the corridor \\
    $Node_{size}\gets$ the current size of all nodes \\
    \begin{algorithmic}[1]
        \Procedure{DeriveCorridor}{$ Node, Node_{old} $}
        \State MoveNode ($ Node $)

        \State $ \Delta x \gets Node.x - Node_{old}.x,~ \Delta y \gets Node.y - Node_{old}.y $

        \State $ slope \gets \frac{\Delta x}{\Delta y}$

        \State $ pLine = $ \Call{DerivePoint}{$ \frac{-1}{ slope } $, $ Node $, $ \mathcal{C}_w $}
        \State $ sideLine_1 = $ \Call{DerivePoint}{$ slope $, $ pLine.a $, $ \mathcal{C}_l $, $ True $}
        \State $ sideLine_2 = $ \Call{DerivePoint}{$ slope $, $ pLine.b $, $ \mathcal{C}_l $, $ True $}
        \State $ corridor \gets
            \begin{bmatrix}
                pLine.a &
                pLine.b \\

                sideLine_1.a &
                sideLine_2.b \\
            \end{bmatrix} $

        \For{$ node^i $ inside $ corridor $}
            \State $ destination = $ \Call{DerivePoint}{$ \frac{-1}{ slope } $, $ node^i $, $ \mathcal{D} $}

            \State MoveNode ($ node^i, destination.b $)
        \EndFor

        \EndProcedure

        \item[]

        \Procedure{DerivePoint}{$ Slope, Point, Length, IsSideLine $}

        \State init. $ a $, $ b $
        \Switch{$Slope$}
            \Case{$0$}
                \If{$ IsSideLine = True $}
                    \State $ a.x \gets Point.x $
                \Else
                    \State $ a.x \gets Point.x + Length $
                \EndIf
                \State $ a.y \gets Point.y $
                \State $ b.x \gets Point.x - Length $,~$ b.y \gets Point.y $
            \EndCase
            \Case{$ \infty $ or $ -\infty $}
                \State $ a.x \gets Point.x $

                \If{$ IsSideLine = True $}
                    \State $ a.y \gets Point.y $
                \Else
                    \State $ a.y \gets Point.y + Length $
                \EndIf

                \State $ b.x \gets Point.x $,~$ b.y \gets Point.y - Length $
            \EndCase
            \Default
                \State $ dx = \sqrt{\frac{Length}{{1+Slope^2}}}$,~ $ dy = Slope \cdot dx $
                \If{$ IsSideLine = True $}
                    \State $ a.x \gets Point.x $,~$ a.y \gets Point.y $
                \Else
                    \State $ a.x \gets Point.x + dx $,~$ a.y \gets Point.y + dy $
                \EndIf

                \State $ b.x \gets Point.x - dx $,~$ b.y \gets Point.y - dy $
            \EndDefault

        \EndSwitch

        \State \Return{$ a, b $}

        \EndProcedure

    \end{algorithmic}
\end{algorithm}

%\end{noindent}

### 2022-02-16 paper/section/related.tex
draft: data
feat: 3 UK rivers with correct sorting
draft: EHR data description
draft: update main section

This section introduces the definition of cartogram types, describes relevant applications of cartograms, and provides a brief overview of some real-world applications of cartogram-based visualizations.

@@

@@

\citea{warf2008Geography} use a Dorling cartogram to represent the religious diversity in the United States. \citea{sun2010Effectiveness} visualize 1996 US election data and 2005 China population date using Dorling, Mosaic, and contiguous cartograms. \citea{cruz2017Adapted} adapts a Dorling cartogram with both contiguous and non-contiguous cartograms to represent the gender pay gap in Portugal. \citea{gao2020Visualising} present a Dorling cartogram to illustrate COVID-19 infections in China. \citea{tong2018cartograms} use a Demers cartogram to visualize health-related date by regions in England, the work introduces a novel technique to remove overlaps of squares based on topological features, aiming to improve both geographical and topological accuracy. \citea{nusrat2020Recognition} investigate the memorability of contiguous and Dorling cartograms by using multiple datasets including demographics, agriculture, and retail data in the US. See Table \ref{table:region vs node} for a list of literature that adopts cartograms for visualization with corresponding geographical regions and node counts.

{
\renewcommand{\arraystretch}{1.5}
\begin{table*}[!tb]
	\centering
	\resizebox{\textwidth}{!}{
		\begin{tabulary}{\textwidth}{|*{4}{l|}r|c|}
			\hhline{~|*{4}{-}|~}
			\multicolumn{1}{c|}{\textbf{Literature}} &
			\textbf{Title} &
			\textbf{Cartogram Type(s)} &
			\textbf{Geographic Region(s)} &
			\textbf{Number of Nodes} &
			\multicolumn{1}{c}{\textbf{Year}} \\
			\hline

		\end{tabulary}
	}
	\caption{}
	\label{table:region vs node}
\end{table*}
}

@@

Cartogram has become a popular choice of visualization in covering various topics by the media, in order to present an unbiased and complete picture of the news. \citea{newman2016Election} generates contiguous cartograms to present US Election results in 2012 and 2016. The Washington Post adopts cartograms to visualize the US overseas economic assistance, arm sales (Mosaic) \cite{bearak2016Everything}, the 2016 US Election (contiguous) \cite{gamio2016Election}, and the Brexit Referendum (Mosaic) \cite{taylor2016What}. The National Geographic uses contiguous and Mosaic cartograms to analyze the 2016 US Election results \cite{miller2016Election}, the same topic is also covered by the Financial Times with a Dorling cartogram \cite{stabe2016Search}. \citea{sandberg2018Cartogram} reports the 2018 US midterm Election with a Mosaic cartogram, the same approach is used to cover the 2020 US Election by the New York Times \cite{thelearningnetwork2020What} and Bloomberg \cite{mccartney20202020}.

>>>>

This section introduces the definition of various cartograms, describes relevant applications of cartograms, and provides a brief overview of some real-world applications of cartogram-based visualizations.

@@

{
	\renewcommand{\arraystretch}{1.5}
	\begin{table*}[!htb]
		\centering
		\resizebox{\textwidth}{!}{
			\begin{tabulary}{\textwidth}{|*{4}{l|}r|c|}
				\hhline{~|*{4}{-}|~}
				\multicolumn{1}{c|}{\textbf{Literature}} &
				\textbf{Title} &
				\textbf{Cartogram Type(s)} &
				\textbf{Geographic Region(s)} &
				\textbf{Number of Nodes} &
				\multicolumn{1}{c}{\textbf{Year}} \\
				\hline

			\end{tabulary}
		}
		\caption{region vs node}
		\label{table:region vs node}
	\end{table*}
}

@@

\citea{warf2008Geography} use a Dorling cartogram to represent the religious diversity in the United States. \citea{sun2010Effectiveness} visualize 1996 US election data and 2005 China population date using Dorling, Mosaic, and contiguous cartograms. \citea{gao2020Visualising} present a Dorling cartogram to illustrate COVID-19 infections in China. \citea{tong2018cartograms} use a Demers cartogram to visualize health-related date by regions in England, the work introduces a novel technique to remove overlaps of squares based on topological features, aiming to improve both geographical and topological accuracy. \citea{nusrat2020Recognition} investigate the memorability of contiguous and Dorling cartograms by using multiple datasets including demographics, agriculture, and retail data in the US. See Table \ref{table:region vs node} for a list of literature that adopts cartograms for visualization with corresponding geographical regions and node counts.

@@

Cartogram has become a popular choice of visualization in covering various topics by the media, in order to present an unbiased and complete picture of the news. The Washington Post adopts cartograms to visualize the US overseas economic assistance, arm sales (Mosaic) \cite{bearak2016Everything}, the 2016 US Election (contiguous) \cite{gamio2016Election}, and the Brexit Referendum (Mosaic) \cite{taylor2016What}. The National Geographic uses contiguous and Mosaic cartograms to analyze the 2016 US Election results \cite{miller2016Election}, the same topic is also covered by the Financial Times with a Dorling cartogram \cite{stabe2016Search}. \cite{cruz2017Adapted} adapts a Dorling cartogram with both contiguous and non-contiguous cartograms to represent the gender pay gap in Portugal. \citea{sandberg2018Cartogram} reports the 2018 US midterm Election with a Mosaic cartogram, the same approach is used to cover the 2020 US Election by the New York Times \cite{thelearningnetwork2020What} and Bloomberg \cite{mccartney20202020}.

% \subsection{Contiguous Applications}

% \citea{newman2016Election} generates contiguous cartograms to present US Election results in 2012 and 2016.

Compared to previous work on Demers Cartogram, we introduce a topological feature, river, aiming to improve the geographical and topological accuracy.
